---
title: "Homework1"
author: "Srikar Murali (#11593114)"
date: "January 30, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(GGally)
```

## Question 1

#### b.

```{r 1b}
x <- matrix(c(1,2,1,-1,3,0,2,7,3,1,-2,-3),
            nrow = 6,
            ncol = 2)
print(x)
mu1 <- colMeans(x)
mu1

cov1 <- cov(x)
cov1

corr1 <- cor(x)
corr1

```

```{r 1c}
colnames(x) <- c('x1', 'x2')


yi1 <- 2*x[,1] + 3
yi1
yi2 <- 3*x[,2] - 2
yi2
corr2 <- cor(yi1, yi2)
corr2

```


## Question 2

#### a.

```{r 2a}

data <- read.csv('C:/Users/srika/Documents/Semester2/STATS 419/Homework1/HW1_data.csv')
data
str(data)
summary(data)
ggpairs(data)
plot(V1 ~ V2, data = data, main = 'V1 vs V2')
plot(V1 ~ V3, data = data, main='V1 vs V3')
plot(V2 ~ V3, data = data, main = 'V2 vs V3')


```

Plot1: As V2 increases V1 seems to increase as well, though not at the same rate, though not at the same rate. The points seem to level off at the V1 value of around 0.85. However there is one outlier which has a value of 0.971 which is far greater than any of the other points.

Plot2: The data seems to be in two distinct clusters. There is still one outlier. V1 seems to increase as V3 increases, though at a slower rate. The points seem to cluster into two main groups.

Plot3: The points seem rather scattered though there seems to be two main groups. There are no significant outliers but there is a divide between the points. There seems to be a relationship between the variables, though it most likely not a linear one.

#### b.

```{r 2b}
mu2 <- colMeans(data)
mu2
cov2 <- cov(data)
cov2

corr2 <- cor(data)
corr2

```


## Question 3

#### c.

```{r 3c}
Z <- t(x)%*%x
Z

Z_inverse <- solve(Z)
Z_inverse

ZinvZ <- Z_inverse%*%Z
ZinvZ

ZZinv <- Z%*%Z_inverse
ZZinv

```

Both of them are approximately equal. No matter the order, multiplying Z by its inverse gives the identity matrix.